---
title: "Maps with R and Leaflet"
author: "Davide Lorino"
date: '2018-07-24'
output:
  html_document:
    df_print: paged
slug: maps-with-r-and-leaflet
tags: []
categories: []
---
Interactive Maps in R with leaflet
2018-07-24

```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE)
```

Plotting interactive maps is easy with the leaflet package in R. In this short guide we’ll use the Launceston Public Seating dataset to plot markers, add pop-ups and freeze the map on an area of interest. Let’s get started.

Load the leaflet and dplyr libraries:

```{r}
library(leaflet)
library(dplyr)
```

Now lets look at our dataset.

## Launceston Seats

On the 20th of July, 2018, the City of Launceston, Tasmania released the ‘Public Seating’ dataset containing 1011 public seats complete with geo-coordinates, materials (aluminium, timber, etc.) and type (seat with back, seat without back, bench, etc.)

The dataset can be found here https://data.gov.au/dataset/public-seating-1c0ab

Import the data into your R environment:

```{r}
seats_data <- read.csv("Public_Seating.csv")
```

Call str() on the dataset and take a look and what we’ve got (i'm ommitting the 13th variable in the dataset because it's an obscurely layered list of metadata - it won't be useful for our purposes):

```{r seats_data}
str(seats_data[,1:12])
```

Just as an aside, you could also do it this way...

```{r}
str(seats_data[,-13])
```

Notice variables X and Y contain the longitude and latitude coordinates of the seats.

Before we plot the seats onto a map, lets look at a basic leaflet.

## Plotting a Leaflet Map

Zoom in and explore the map - can you find your house?

```{r library(leaflet)}
leaflet() %>%
  addTiles
```

Looks good, but it’s not really focused on Tasmania - when we plot our seats we’ll have to navigate manually in the map in order to find them - not ideal. 

Let’s make it so the map looks directly at Launceston, Tasmania from the get-go.

```{r}
leaflet() %>%
  addTiles() %>%
  setView(lng = 147.1376, lat = -41.43584, zoom = 12)
```

Cool, now we can restrict the movement of the map so that it doesn’t let you stray too far by accident.

```{r}
leaflet(options = leafletOptions(
  minZoom = 5, dragging = TRUE)) %>%
  addTiles() %>%
  setView(lng = median(seats_data$X), lat = median(seats_data$Y),
            zoom = 13) %>%
  setMaxBounds(lng = min(seats_data$X) - .05, lat = min(seats_data$Y) - 0.5,
               lng2 = max(seats_data$X) + .05, lat2 = max(seats_data$Y) + .05)
```

In the map above, try to zoom out and move away from Tasmania - the map pulls back.

Now let’s add markers for each seat in the dataset, pay close attention to the line that starts with ‘addMarkers’:

```{r}
leaflet(options = leafletOptions(
  minZoom = 5, dragging = TRUE)) %>%
  addTiles() %>%
  addMarkers(lng = unique(seats_data$X), lat = unique(seats_data$Y)) %>%
  setView(lng = median(seats_data$X), lat = median(seats_data$Y),
          zoom = 13) %>%
  setMaxBounds(lng = min(seats_data$X) - .05, lat = min(seats_data$Y) - 0.5,
               lng2 = max(seats_data$X) + .05, lat2 = max(seats_data$Y) + .05)
```

Ok great, now lets add popups so that the type of seat shows up when we click a marker:

```{r}
leaflet(options = leafletOptions(
  minZoom = 5, dragging = TRUE)) %>%
  addTiles() %>%
  addMarkers(lng = unique(seats_data$X), lat = unique(seats_data$Y), popup = seats_data$class) %>%
  setView(lng = median(seats_data$X), lat = median(seats_data$Y),
          zoom = 13) %>%
  setMaxBounds(lng = min(seats_data$X) - .05, lat = min(seats_data$Y) - 0.5,
               lng2 = max(seats_data$X) + .05, lat2 = max(seats_data$Y) + .05)
```


And there we go! Maps are easy with R and leaflet.

` ——————————————————